---
title: "sleep-medicine-review-1"
author: "Tom Walton"
date: "2024-09-13"
output: html_document
bibliography: review-1.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Reviewer 1

The aim of this study is to identify and evaluate the methodology that has been used for conceptualizing sleep discrepancy, and the authors try to focus on the variety of methodology itself rather than clinical significance of sleep misperception and mismeasurement. For this purpose, the authors made the use of various search engines and have shown a thorough list of previous studies. Therefore, this list has a significant value for the researchers who are tackling this theme, but the journal review is not an encyclopedia, and this manuscript is difficult to follow.
Although sleep discrepancy is not limited in the area of insomnia research, the concept of sleep discrepancy originates from sleep state misperception or paradoxical insomnia. I understand the authors' viewpoint described in 5.10 (The scope of sleep discrepancy research is likely to have been underestimated), however, the whole presentation would have been more organized if they restricted the references to those relating to normative/insomnia data. I guess, for the majority of readers, focus of interest is about the methodology used for this population, and it will be the next step to widen this field.
Another aspect which made this manuscript difficult to follow is that the authors mixed the two objective methods to measure sleep, one is PSG, and the other actigraphy. The situation of utilizing these two methods is different, which relates to the purpose of each study, and the authors might have made it more clear by the terms of episodic sleep and habitual sleep. From this aspect, I suggest the authors should divide the result section, one for PSG, and the other for actigraphy. In addition, it may be useful if the authors put a table showing the combination of the plural objective and subjective methods.

# Response to Reviewer 1
We appreciate the comments about the breadth of the topic making the manuscript difficult to follow. The amount of information contained within the review is vast and needs to be digestible if it is to be useful to the reader. We are concerned that restricting the results to investigations of normative and insomnia populations would not appreciably reduce the complexity of the results section as the length here is due mainly to the comprehensiveness of recorded methodological features. To enhance readability we have added a hierarchical structure to the results, breaking methodological features into the stages of measurement, data preprocessing, statistical analysis, and study design. We hope that this serves to organise the information better and communicate the research more effectively. 
In relation to the mixing of PSG and actigraphy, we agree the purposes of using these methods are often separable however the distinction is not always perfect. For instance, whilst PSG is often recorded for one night or two nights, sometimes multiple nights are used. Sometimes another EEG-based method or ambulatory system is used in a naturalistic environment. I believe there were even some cases of using actigraphy for single night sleep esimates. Interrogating the data just now, there appear to be six such examples:

``` {r overlap}

# load packages
library("tidyverse")

# Read data set into data frame
full_screen <- read.csv("../data/review_data.csv")

# Trim excluded studies from the data frame
screen <- filter(full_screen, fulltext_include=="Yes" | fulltext_include=="Second study")

# Count single night studies using actigraphy but not PSG (some studies use both)
length(which(screen["obj_acti"] == 1 & is.na(screen["obj_PSG"]) & screen["rm_singlenight"] == 1))
```

It would be very difficult to organise or split the result sections according to any of the recorded methodological features due to the ubiquity of this kind of overlap.
Thank you for the suggestion of adding a table of plural features. We have added a list of all recorded methodological features to the text which may be compared to the data dictionary and full table available on github if readers are interested in consulting the data directly.

# Reviewer 2

The authors offer a valuable review focused on sleep discrepancy. The material is well-organized and clearly illustrated and typos are rarely identified, i.e., Sleep discrepancy has been investigated with diverse methods for its conceptualisation and measurement, such that it can be difficult (to?) integrate findings across studies….An individual is unable to directly estimate their (?) number of EEG arousals, quantity or proportion of N3 sleep, or other features of sleep macro or microstructure unavailable to consciousness.
However, authors are warmly invited to clarify their viewpoint on the definition of sleep discrepancy. One sentence states: Sleep discrepancy is a common feature of insomnia disorder, where it is also referred to as sleep misperception or paradoxical insomnia. But a few lines after we find:….paradoxical insomnia, an insomnia subvariant defined, in part, by the presence of sleep discrepancy.
I don't want to criticize the adopted methodological selection, but in my opinion the mismatch between subjective experience and objective measures is not limited to insomnia. We daily interview patients with sleep apnea, periodic myoclonus or RBD who show impressive alterations in conventional sleep measures but not always complain of poor sleep quality or excessive daytime sleepiness.
In an ideal world:
* If you subjectively experience a short sleep duration, your objective TST and SE should be reduced.
* If you subjectively experience several awakenings, your objective WASO should be enhanced.
* If you subjectively experience a superficial sleep, your objective N3 should be attenuated.
* If you subjectively experience a non-restorative sleep, your objective sleep metrics could be apparently normal but are commonly accompanied by increased amounts of EEG arousals and cyclic alternating pattern (CAP). This is what happens in paradoxical insomnia (Parrino et al. 2009).
When this association occurs, we exploit the major potential benefits of both PSG macro- and microstructural information (this means that time is ripe to include CAP metrics in the conventional scoring activity).
Unfortunately, a certain percentage of patients with insomnia, parasomnia, OSA, PLM, still show a mismatch between subjective and objective measurements suggesting that the official classifications of sleep disorders and sleep scoring may undergo impressive and embarrassing leakages.
Perhaps, other biomarkers are neglected (REM sleep fragmentation? autonomic imbalance? mental constraints?) or, more likely, sleep quality resembles a cake that requires balanced amounts of multiple ingredients (milk, butter, flour, sugar, yeast) to become a delicious pie?
Out of metaphor, Efficiency, Continuity, Depth, Stability, Regional EEG processes, Vegetative Burden, Cognitive Modulation are probably distinct ingredients of the sleep quality recipe, which are waiting to be blended in a comprehensive (mathematical) formula aiming to fill the gap between subjective and objective realms.
Maybe we will never reach a 100% overlap, but a review on sleep discrepancy can be an excellent opportunity to face these challenging issues: we hope the authors will not miss the chance.

# Response to Reviewer 2

Thank you for identifying the typos and the apparently contradictory introduction of paradoxical insomnia. We have corrected the former and clarified our language with the latter.
We wholeheartedly agree with your perspective on the importance of sleep discrepancy outside of insomnia.
<!-- continue to summarise the perspective -->  Your descriptions are comparable to our own clinical experiences and...

In our view, your critique may apply more to our inclusion criteria rather than the adopted methodological selection. The breadth of the measured characteristics reflects the fact that there was very little selection going on--we ended up including almost everything. By including only studies that directly compared equivalent measures of self-report and objective sleep, studies that measured the many possible metrics and correlates of sleep quality and finer-grained objective characteristics of sleep were likely sometimes omitted. The presence of discrepancy between objective sleep quality and sleep complaint is certainly interesting and of profound clinical importance but ended up lying a little outside of the scope of the review. We are unsure if sleep discrepancy is even the best term for this phenomenon. @lichstein2017 suggested "uncoupled sleeping" but we aren't sure about the use of this term either. At any rate, the discussion section of the present review should situate itself within this context and we agree that it would be a shame to miss the chance to do this at the very least. We have changed the discussion to incoporate this, particularly in the sections discussing sleep quality discrepancy and the conceptual differences between forms of sleep discrepancy.

# Summary of changes to manuscript

1. Results have been organised hierarchically according to their stage in the research process
2. A list of recorded methodological features has been added to the text
3. The discussion has been modified to better reflect the current clinical context of sleep discrepancy as a concept

# References

